### 2019-11-11 tesis_hiram/Chapter1/chapter1.tex
final
tesis-arreglo1

Particle physics is a field of study of physics that studies the particles that create matter and its interactions to very small scale, even smaller than atoms. The particles analized in particle physics are the fundamental particles that compose the matter and the responsibles for all forces of nature. All thoses particles and their properties are grouped in a theory called the Standard Model. Standard Model (SM) is a theory that describes the fundamental forces (except gravity) and classify the elementary particles in bosons (entire spin) and fermions (half spin).
The SM, created almost 50 years ago, has been the most successul theory in explain the phenomena in the subatomic world.
\\

One of the particles of the SM, the Higgs Boson, was recently discovered in 2012 in the ATLAS and CMS experiments at CERN. The discovery of Higgs boson in proton proton collisions marked a new era in the particle physics due to its characteristic of giving the matter the property of mass, according to previous theories and predictions. Since then, people around the world are working in the Large Hadron Collider at CERN in order to detect the Higgs bosons in higher energies and differents Higgs production processes. 
Also the impact of Higgs is also important in cosmology, where theoretical investigations relate the Higgs boson and the cosmic inflation.\\

In the recent years, there have been many studies in the Higgs production processes at CERN in pp collisions, such as gluon fusion (ggF), VBF, WH, ZH and $t\bar{t}$H and people have obtained consistent results in the search of Higgs in those processes. 
In this work, we will analize the production of Higgs boson in association with a
single top quark (tH) in proton-proton collisions with the CMS experiment of the LHC. 
The motivation of the work is this mechanism of production of the Higgs boson has not been observed before by any experiment. The exploration of Higgs production on the tH channel is subject relatively new. %The SM is used to study the tH channel and compare it with other Higgs production mechanisms results.
\\
Many channels are yet out of
reach experimentally. At the same time, many theoretical
proposals are yet to be put on test. It is possible that studies on the Higgs boson give evidence of small deviations from the SM and test new physics that are beyond SM such as String Theory and Supersymmetry. \\
%********************************** % Third Section *************************************

@@

The last group of the family are the bosons,which have a spin 1 and 0. 
There are 5 bosons: $W^{\pm}$, $Z$ with neutral charge are responsibles for weak force; photon, a massless particle, is responsible for electromagnetic force; gluon which also massless, is related to strong force and Higgs boson, recently discovered, is in charge of give the property of mass to particles via Higgs mechanism. Bosons let other particles to interact, even themselves, which is shown in the figure \ref{sm}. 
\\
The $W$,$Z$ and Higgs bosons, can interact with the most of the particles, while the gluon can only interact with quarks and itself.
Those bosons act as mediators in decays and creation of complex
particles, as mesons and baryons.\\

@@

There are experiments that detects particles around the world, such as the large hadron collider (LHC) in Europe, the Fermilab in United States of America, KEK in Japan, etc. More information about the LHC will be described in the next chapters.
\\
Since quarks and gluons only exist in bound states, it is imposible to observe isolated quarks except the top quark, which decays before it has time to form a
bound state. As two quarks are separated, the strong force between them increases
which is equivalent to the potential energy increasing. This energy transforms into
quark-anti-quark pairs. \\

@@

The production of top quark can be generated in differents ways. In tevatron and LHC, the production of $t\bar{t}$ comes from ggF (gluon gluon fusion) via strong force. Even it is posible to create a single top quark in the LHC via electroweak interactions (with a W boson)\cite{th1}. 
The top quark, due to its big mass, always decays into a W boson and a b quark. But since the W boson have various decays, the top quark can generate many posible particles, acording to W boson decays. Some posible decays of top quark is the decay to a lepton such as muon, electron or tau and their neutrinos. %The antitop, antiparticle of top quark, can decay to W$^{-1} \bar{b}$ and its decay is the same as top, but with inverse signal.

@@

\section{Higgs production mechanisms}
During a particle collision, there are various ways that the Higgs boson can be created. The most common is the gluon gluon fusion process ($gg$F) which is the process which involve two gluons mediated by the exchange of a virtual, heavy top
quarks. Another process is vector boson fusion (VBF), where two fermions interact via a vector boson ($W$ or $Z$ ) and create a Higgs along with other particles. There is also the vector higgs process (VH), where the creation of particles using vector boson comes from a interaction between fermions and anti fermions\cite{pd}. These processes are shown in the figure \ref{psu} \\

@@

The number of particles decreases over time exponentially and this can be measured. With this, it is possible determinate the mean lifetime\cite{griff}
\begin{align}
\tau=\frac{1}{\Gamma}
\end{align}
In case of the decay rate to a specific process, it is necessary get the branching fraction of the process. 
The branching ratio for a decay process is the ratio of the number of particles which decay via a specific decay mode with respect to the total number of particles which decay via all decay modes.
\begin{align}
BR_i =\frac{\Gamma_i}{\sum_{i}\Gamma_i}
\end{align}
Where $\Gamma=\sum_i\Gamma_i$ is the total decay width (sum of all partial widths) of the particle and is related to lifetime of the particle: $\Gamma=1/\tau$.
Since the dimension of $\Gamma$ is the inverse of time, in our system of natural units, it is measured in inverse seconds,it has the same dimension as mass (or energy)\footnote{Cleaves H.J. (2011) Branching Ratio. In: Gargaud M. et al. (eds) Encyclopedia of Astrobiology. Springer, Berlin, Heidelberg}. The lifetime of the Higgs boson is predicted to be 1.56 $\times$ $10^{-22}$ seconds and corresponds to $\Gamma_H$ is about 4 MeV, but has not yet been measured due to the detector resolution\cite{cms-manual}.
\\
\begin{table}[!htbp] 
\caption[SM Higgs boson branching ratios for $M_H$ =125 GeV]{SM Higgs boson branching ratios for $M_H$ =125 GeV \protect \cite{pd}}
\centering
\begin{tabular}{|c|c|}
\hline
Higgs decay & Branching ratio (BR)\\
\hline
$H \rightarrow$ b$\bar{b}$ &$58.4\%$ \\
\hline
 $H \rightarrow$ $W^+W^-$ &$21.4\%$ \\
\hline
$H \rightarrow$ $\tau^+ \tau^-$ & $6.27\%$\\
\hline
$H \rightarrow$ ZZ &$2.62\%$\\
\hline
$H \rightarrow$ $\gamma\gamma$ &$0.227\%$\\
\hline
$H \rightarrow$ Z$\gamma$ &$0.153\%$\\
\hline
$H \rightarrow$ $\mu^+\mu^-$ &$0.0218\%$\\
\hline
\end{tabular}
\label{higgs1}
\end{table}
The decay $H \rightarrow$ $b\bar{b}$ have a branching ratio a bit more than 50$\%$, therefore is the most likely decay Higgs boson can decay. Due to this, $H\rightarrow b\bar{b}$ is a very studied process and most likely to detect a Higgs boson in the data after reconstruction. The decay $H \rightarrow$ $\mu^+\mu^-$ has a very low branching ratio. This decay is very rare but not imposible to detect in the future. In this work the majority of the Higgs decays detected are from $WW$, $ZZ$, $\tau \tau$\\

>>>>

Particle physics is a field of study of physics that studies the particles that create matter and its interactions to very small scale, even smaller than atoms. The particles analyzed in particle physics are the fundamental particles that compose the matter and the responsible for all forces of nature. All those particles and their properties are grouped in a theory called the Standard Model. Standard Model (SM) is a theory that describes the fundamental forces (except gravity) and classify the elementary particles in bosons (entire spin) and fermions (half spin).
The SM, created almost 50 years ago, has been the most successful theory in explain the phenomena in the subatomic world.
\\

One of the particles of the SM, the Higgs Boson, was recently discovered in 2012 in the ATLAS and CMS experiments at CERN. The discovery of Higgs boson in proton proton collisions marked a new era in the particle physics due to its characteristic of giving the matter the property of mass, according to previous theories and predictions. Since then, people around the world are working in the Large Hadron Collider at CERN in order to detect the Higgs bosons in higher energies and different Higgs production processes. 
Also the impact of Higgs is also important in cosmology, where theoretical investigations relate the Higgs boson and the cosmic inflation.\\

In the recent years, there have been many studies in the Higgs production processes at CERN in pp collisions, such as gluon fusion (ggF), VBF, WH, ZH and $t\bar{t}$H and people have obtained consistent results in the search of Higgs in those processes. 
In this work, we will analyze the production of Higgs boson in association with a
single top quark (tH) in proton-proton collisions with the CMS experiment of the LHC. 
The motivation of the work is this mechanism of production of the Higgs boson has not been observed before by any experiment. The exploration of Higgs production on the tH channel is subject relatively new. %The SM is used to study the tH channel and compare it with other Higgs production mechanisms results.
\\
Many channels are yet out of
reach experimentally. At the same time, many theoretical
proposals are yet to be put on test. It is possible that studies on the Higgs boson give evidence of small deviations from the SM and test new physics that are beyond SM such as String Theory and Supersymmetry. \\
%********************************** % Third Section *************************************

@@

The last group of the family are the bosons,which have a spin 1 and 0. 
There are 5 bosons: $W^{\pm}$, $Z$ with neutral charge are responsible for weak force; photon, a massless particle, is responsible for electromagnetic force; gluon which also massless, is related to strong force and Higgs boson, recently discovered, is in charge of give the property of mass to particles via Higgs mechanism. Bosons let other particles to interact, even themselves, which is shown in the figure \ref{sm}. 
\\
The $W$,$Z$ and Higgs bosons, can interact with the most of the particles, while the gluon can only interact with quarks and itself.
Those bosons act as mediators in decays and creation of complex
particles, as mesons and baryons.\\

@@

There are experiments that detects particles around the world, such as the large hadron collider (LHC) in Europe, the Fermilab in United States of America, KEK in Japan, etc. More information about the LHC will be described in the next chapters.
\\
Since quarks and gluons only exist in bound states, it is impossible to observe isolated quarks except the top quark, which decays before it has time to form a
bound state. As two quarks are separated, the strong force between them increases
which is equivalent to the potential energy increasing. This energy transforms into
quark-anti-quark pairs. \\

@@

The production of top quark can be generated in different ways. In tevatron and LHC, the production of $t\bar{t}$ comes from ggF (gluon gluon fusion) via strong force. Even it is possible to create a single top quark in the LHC via electroweak interactions (with a W boson)\cite{th1}. 
The top quark, due to its big mass, always decays into a W boson and a b quark. But since the W boson have various decays, the top quark can generate many possible particles, according to W boson decays. Some posible decays of top quark is the decay to a lepton such as muon, electron or tau and their neutrinos. %The antitop, antiparticle of top quark, can decay to W$^{-1} \bar{b}$ and its decay is the same as top, but with inverse signal.

@@

\section{Higgs production mechanisms}
During a particle collision, there are various ways that the Higgs boson can be created. The most common is the gluon gluon fusion process ($gg$F) which is the process which involve two gluons mediated by the exchange of a virtual, heavy top
quarks. Another process is vector boson fusion (VBF), where two fermions interact via a vector boson ($W$ or $Z$ ) and create a Higgs along with other particles. There is also the vector Higgs process (VH), where the creation of particles using vector boson comes from a interaction between fermions and anti fermions\cite{pd}. These processes are shown in the figure \ref{psu} \\

@@

The number of particles decreases over time exponentially and this can be measured. With this, it is possible determinate the mean lifetime\cite{griff}
\begin{align}
\tau=\frac{1}{\Gamma}
\end{align}
In case of the decay rate to a specific process, it is necessary get the branching fraction of the process. 
The branching ratio for a decay process is the ratio of the number of particles which decay via a specific decay mode with respect to the total number of particles which decay via all decay modes.
\begin{align}
BR_i =\frac{\Gamma_i}{\sum_{i}\Gamma_i}
\end{align}
Where $\Gamma=\sum_i\Gamma_i$ is the total decay width (sum of all partial widths) of the particle and is related to lifetime of the particle: $\Gamma=1/\tau$.
Since the dimension of $\Gamma$ is the inverse of time, in our system of natural units, it is measured in inverse seconds,it has the same dimension as mass (or energy)\footnote{Cleaves H.J. (2011) Branching Ratio. In: Gargaud M. et al. (eds) Encyclopedia of Astrobiology. Springer, Berlin, Heidelberg}. The lifetime of the Higgs boson is predicted to be 1.56 $\times$ $10^{-22}$ seconds and corresponds to $\Gamma_H$ is about 4 MeV, but has not yet been measured due to the detector resolution\cite{cms-manual}.
\\
\begin{table}[!htbp] 
\caption[SM Higgs boson branching ratios for $M_H$ =125 GeV]{SM Higgs boson branching ratios for $M_H$ =125 GeV \protect \cite{pd}}
\centering
\begin{tabular}{|c|c|}
\hline
Higgs decay & Branching ratio (BR)\\
\hline
$H \rightarrow$ b$\bar{b}$ &$58.4\%$ \\
\hline
 $H \rightarrow$ $W^+W^-$ &$21.4\%$ \\
\hline
$H \rightarrow$ $\tau^+ \tau^-$ & $6.27\%$\\
\hline
$H \rightarrow$ ZZ &$2.62\%$\\
\hline
$H \rightarrow$ $\gamma\gamma$ &$0.227\%$\\
\hline
$H \rightarrow$ Z$\gamma$ &$0.153\%$\\
\hline
$H \rightarrow$ $\mu^+\mu^-$ &$0.0218\%$\\
\hline
\end{tabular}
\label{higgs1}
\end{table}
The decay $H \rightarrow$ $b\bar{b}$ have a branching ratio a bit more than 50$\%$, therefore is the most likely decay Higgs boson can decay. Due to this, $H\rightarrow b\bar{b}$ is a very studied process and most likely to detect a Higgs boson in the data after reconstruction. The decay $H \rightarrow$ $\mu^+\mu^-$ has a very low branching ratio. This decay is very rare but not impossible to detect in the future. In this work the majority of the Higgs decays detected are from $WW$, $ZZ$, $\tau \tau$\\

### 2019-11-11 tesis_hiram/Chapter2/chapter2.tex
tesis-arreglo1

The main objetive of the LHC is to study the nature of electroweak symmetry breaking for the Higgs mechanism of the SM. 
Alternatives to the Standard Model that involves different symmetries are also put on test. With the exploration of theses theories, people hope for a discovery that guide them toward a unified theory, so the importance to accelerate particles to higher energy scales.
\\
The LHC has a schedule where there is a period of experiments that last for three or four years. Ending the experimental time, the engineers of the LHC start the maintenance.

@@

\section{The Compact Muon Solenoid}	
The Compact Muon Solenoid (CMS) is a detector with multiple uses in the LHC and part of the main experiments at CERN. Located underground in the France- Switzerland border, in the city of Cessy, France.
This detector was designed in the early 1990s, based on the mass limit of the Higgs boson, and put on operation in 2008, it has a big solenoid that generates a great magnetic field of 4 teslas with the objective to separate particles after a particle collision.
 The detector is 21 metres long, 15 meters wide, 15 meters high, it has a diameter of 5.9 m with a weight of 12000 ton. The reason for such a strong magnetic field is to obtain a better momentum resolution.
 The magnetic flux is returned via a 1.5 m thick iron yoke instrumented with four stations of muon
 chambers.

@@

The Strip Detector have almost 15 400 modules, mounted on
carbon-fibre structures and housed inside by a temperature controlled outer support tube. The operating temperature must be around -$\ang{20}$C. 
The total area of the pixel detector is around 1 $m^2$ and the silicon strips is 200 m$^2$. 
The inner tracker comprises 66 million pixels and 9.6
million silicon strips \cite{cms-manual}.
A full coverage of the Tracker is shown in figure \ref{pixel2}
\\

@@

The electromagnetic calorimeter or ECAL is a calorimeter composed of 61200 lead tungstate (PbWO$_4$) crystals that act as scintillanting crystals (emits light when particles interact with the crystal). These calorimeters are mounted in the central barrel and closed by 7324 crystals in the both endcaps. The barrel section of the ECAL covers the range $|\eta| < 1.479$. The crystals in the barrel have a pyramid shape and their crosss section is 22$\times$22 mm$^2$ at the front face and 26$\times$26 mm$^2$ at rear face. The crystal length is 230 mm. The EB (barrel section of ECAL) has a inner radius of 129 cm\cite{cms-manual}\\

@@

\subsection{Hadron calorimeter}
The hadron calorimeter (HCAL) along the ECAL, form a complete calorimeter that allows to measure the jets and missing transverse energy. HCAL is located in the barrel and the endcaps, surrounding the ECAL and affected by the magnetic field generated by the solenoid. The Barrel section (HB) and endcap section (HE) cover the pseudorapity 0 < $|\eta|$ < 1.3 and 1.3 < $|\eta|$ < 3.0 respectively.\\

@@

HE is composed by brass absorber plates, with the thickness of 78 mm, and a total of 19 scintillator layers with a thickness of 3.7 mm. HE is sectioned in 5 in $\phi$ to match the barrel wedges.
. %HO scintillators follow the HCAL barrel tower geometry in $\eta$ and angle $\phi$, that is, the geometry of the barrel muon system. Divided by 5 rings of 2.54 wide, composed of the scintillators on either side of the tail catcher iron of 18cm thick at radial distances of 3850 and 4097 mm respectively. \cite{cms-manual} 
\\
\begin{figure}[!htbp]
	\centering
	\includegraphics[width=10cm,height=6cm]{Chapter2/hcal.png}
	\caption{View of the CMS HCAL showing the different layers and tower regions. In the endcap, the towers are defined with two longitudinal segments as shown by the colors\cite{cms-manual}}\label{hcal}
\end{figure}
The outer barrel hadron calorimeter (HO) consists of layers of scintillator with thickness of 10 mm, located outside of the
magnet coil that cover the region -1.26 < $|\eta|$ < 1.26.\\
There is also a foward calorimeter (HF) located at 11.2 m from the interaction point, that cover the region of 2.9 < $|\eta|$ < 5. They are made of
steel absorbers and embedded radiation hard quartz fibers, which collects 
Cherenkov light\cite{cms-manual}.% that is when a charged particle travels through matter faster than light can, resulting in a blue light emmision of the material. 
%The length of the fibers are 1.65 m and 1.43 m, which are alterned with a 5 mm of separation. 
\\

>>>>

The main objective of the LHC is to study the nature of electroweak symmetry breaking for the Higgs mechanism of the SM. 
Alternatives to the Standard Model that involves different symmetries are also put on test. With the exploration of theses theories, people hope for a discovery that guide them toward a unified theory, so the importance to accelerate particles to higher energy scales.
\\
The LHC has a schedule where there is a period of experiments that last for three or four years. Ending the experimental time, the engineers of the LHC start the maintenance.

@@

\section{The Compact Muon Solenoid}	
The Compact Muon Solenoid (CMS) is a detector with multiple uses in the LHC and part of the main experiments at CERN. Located underground in the France- Switzerland border, in the city of Cessy, France.
This detector was designed in the early 1990s, based on the mass limit of the Higgs boson, and put on operation in 2008, it has a big solenoid that generates a great magnetic field of 4 teslas with the objective to separate particles after a particle collision.
 The detector is 21 meters long, 15 meters wide, 15 meters high, it has a diameter of 5.9 m with a weight of 12000 ton. The reason for such a strong magnetic field is to obtain a better momentum resolution.
 The magnetic flux is returned via a 1.5 m thick iron yoke instrumented with four stations of muon
 chambers.

@@

The Strip Detector have almost 15 400 modules, mounted on
carbon-fiber structures and housed inside by a temperature controlled outer support tube. The operating temperature must be around -$\ang{20}$C. 
The total area of the pixel detector is around 1 $m^2$ and the silicon strips is 200 m$^2$. 
The inner tracker comprises 66 million pixels and 9.6
million silicon strips \cite{cms-manual}.
A full coverage of the Tracker is shown in figure \ref{pixel2}
\\

@@

The electromagnetic calorimeter or ECAL is a calorimeter composed of 61200 lead tungstate (PbWO$_4$) crystals that act as scintillating crystals (emits light when particles interact with the crystal). These calorimeters are mounted in the central barrel and closed by 7324 crystals in the both endcaps. The barrel section of the ECAL covers the range $|\eta| < 1.479$. The crystals in the barrel have a pyramid shape and their cross section is 22$\times$22 mm$^2$ at the front face and 26$\times$26 mm$^2$ at rear face. The crystal length is 230 mm. The EB (barrel section of ECAL) has a inner radius of 129 cm\cite{cms-manual}\\

@@

\subsection{Hadron calorimeter}
The hadron calorimeter (HCAL) along the ECAL, form a complete calorimeter that allows to measure the jets and missing transverse energy. HCAL is located in the barrel and the endcaps, surrounding the ECAL and affected by the magnetic field generated by the solenoid. The Barrel section (HB) and endcap section (HE) cover the pseudorapidity 0 < $|\eta|$ < 1.3 and 1.3 < $|\eta|$ < 3.0 respectively.\\

@@

HE is composed by brass absorber plates, with the thickness of 78 mm, and a total of 19 scintillator layers with a thickness of 3.7 mm. HE is sectioned in 5 in $\phi$ to match the barrel wedges.
. %HO scintillators follow the HCAL barrel tower geometry in $\eta$ and angle $\phi$, that is, the geometry of the barrel muon system. Divided by 5 rings of 2.54 wide, composed of the scintillators on either side of the tail catcher iron of 18cm thick at radial distances of 3850 and 4097 mm respectively. \cite{cms-manual} 
\\
\begin{figure}[!htbp]
	\centering
	\includegraphics[width=10cm,height=6cm]{Chapter2/hcal.png}
	\caption{View of the CMS HCAL showing the different layers and tower regions. In the endcap, the towers are defined with two longitudinal segments as shown by the colors\cite{cms-manual}}\label{hcal}
\end{figure}
The outer barrel hadron calorimeter (HO) consists of layers of scintillator with thickness of 10 mm, located outside of the
magnet coil that cover the region -1.26 < $|\eta|$ < 1.26.\\
There is also a foward calorimeter (HF) located at 11.2 m from the interaction point, that cover the region of 2.9 < $|\eta|$ < 5. They are made of
steel absorber and embedded radiation hard quartz fibers, which collects 
Cherenkov light\cite{cms-manual}.% that is when a charged particle travels through matter faster than light can, resulting in a blue light emmision of the material. 
%The length of the fibers are 1.65 m and 1.43 m, which are alterned with a 5 mm of separation. 
\\
